---
title: "Risk Management Project"
author: Chloe Morin-Leclerc, Felix-Antoine Groulx, Denis Genest, Thien Duy Tran
date: "`r format(Sys.time(), '%B %e, %Y')`"
output: html_notebook
---

<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

# Context
You work as a quantitative analyst for a large investment bank. You and your team are responsible for
challenging the models used by traders and risk managers. You work with R and love reproducible research.
All your files are written in Rmarkdown or R notebook.

You can watch the introductory video on Rmarkdown to help you build properly the R file.
Moreover, you use GitHub with your team. You’ll build a dedicated project for the tasks below and use
RStudio with GitHub extensively. You also use Loom to share your findings with other teams in the investment
bank.

## Learning objectives
Content (scientific rigor, concepts, creativity)\
• Choose the right tools.\
• Implement the steps correctly.\
• Come up with innovative solutions.\

## Form (coding, collaboration, presentation)
• Build RStudio project with proper folder structure and Rmarkdown/nootebook file to reproduce your results.\
• Program with state-of-the-art coding standards.\
• Use GitHub repository for collabortive research.\
• Use Loom video for presenting your results.

# Project (Risk management)
The objective is to implement (part of) the risk management framework for estimating the risk of a book of
European call options by taking into account the risk drivers such as underlying and implied volatility.

## Instructions

## Data
Load the database Data. Identify the price of the S&P 500, the VIX index, the term structure of
interest rates (current and past), and the traded options (calls and puts).

```{r}
# Packages

# install.packages("here")
library("here")

# load data
load(file = here("Data", "Market.rda"))

# load function 
source(file = here("Functions", "price_call.r")) # function that price Calls with Black-Shcole  formula
source(file = here("Functions", "lin_inter.r"))  # interpolation lineaire entre les taux d'interet de la courbe d'interet

# data
vix    <- Market$vix
sp_500 <- Market$sp500
rf     <- Market$rf
calls  <- Market$calls
puts   <- Market$puts

```

## Pricing of a portfolio of options
Assume the following book of European call options: 1x strike K = 1600 maturity 20 days, 1x strike
K = 1650 maturity 20 days, 1x strike K = 1750 maturity 40 days and 1x strike K = 1800 maturity 40
days.

```{r}
# Portfolio  Calls - Caracteristics
book <- matrix(data=NA, nrow = 4, ncol = 4)
colnames(book) <- c("Quantity","Call","StrikePrice","Maturity")

book[1,] <- c(1, 1, 1600, 20 / 365)
book[2,] <- c(1, 1, 1650, 20 / 365)
book[3,] <- c(1, 1, 1750, 40 / 365)
book[4,] <- c(1, 1, 1800, 40 / 365)

# input for black-Schole calls pricing 
# interest rate interpolated
r_1 <- lin_inter(book[1,4])
r_2 <- lin_inter(book[3,4])
# Spot price, Volatility, Strike price, Maturity
S   <- sp_500[length(sp_500)]
vol <- vix[length(vix)]
K   <- book[,3]
N   <- book[,4] 

# calculating and Stocking Call's prices
call_price <- c(rep(NA, 4))
call_price[1] <- Price_call(S, K[1], r_1, vol, N[1])
call_price[2] <- Price_call(S, K[2], r_1, vol, N[2])
call_price[3] <- Price_call(S, K[3], r_2, vol, N[3])
call_price[4] <- Price_call(S, K[4], r_2, vol, N[4])

# stocking Portfolio Value
PF_price <- sum(call_price)
```

Find the price of this book given the latest underlying price and the latest implied volatility (take the
VIX for all options). Use Black-Scholes formula to price the options. Take the current term structure
and linearly interpolate to find the corresponding rates. Use 360 days/year for the term structure and
250 days/year for the time to maturity of the options.


```{r}
# log return - sp_500
log_return <-diff(log(sp_500))

# model calibration
tmp    <- length(log_return)

mu_hat <- mean(log_return[seq(1,tmp,5)])
s2_hat <- mean((log_return[seq(1,tmp,5)] - mu_hat)^2) #n/(n-1)
s_hat  <- s2_hat^0.5

# number of simulation
H <- 10000
t <- 5 / 365

# Portfolio  Calls - Caracteristics - 5 days later
book_2 <- book                 # same as previous except days until maturity
book_2[,4] <- book_2[,4] - t   # t days later

# log return sp_500 projected in 5 days
# set seed(1234).....
sample <- rnorm(H, mean=mu_hat, sd=s_hat)

# price sp_500 projected in 5 days
S_2 <- sp_500[length(sp_500)] * exp(sample)

# input for black-Schole calls pricing
r_3   <- lin_inter(book_2[1,4])
r_4   <- lin_inter(book_2[3,4])
vol_2 <- vix[length(vix)]
K_2   <- book_2[,3]
N_2   <- book_2[,4] 

# stocking matrix (Hx4) - H simulation, 4 calls 
call_price_2 <- matrix(NA, nrow = H, ncol = 4)

# calculating and stocking call's price simulation
for (i in 1:10000){
  call_price_2[i,1] <- Price_call(S_2[i], K_2[1], r_3, vol_2, N_2[1])
  call_price_2[i,2] <- Price_call(S_2[i], K_2[2], r_3, vol_2, N_2[2])
  call_price_2[i,3] <- Price_call(S_2[i], K_2[3], r_4, vol_2, N_2[3])
  call_price_2[i,4] <- Price_call(S_2[i], K_2[4], r_4, vol_2, N_2[4])
}

# stocking H portfolio value 
PF_price_2 <- apply(call_price_2,1,sum)

# P&N of the H simulation 
r_pl <- lin_inter(t)  #actualisation factor
pl   <- PF_price_2 * exp(-t *r_pl) - PF_price

# VaR and EVaR
alpha <- 0.95
VaR <- sort(pl)[(1 - alpha) * H]
EVaR <- mean(sort(pl)[1:((1 - alpha) * H)])

# histogram
hist(pl, nclass = round(10 * log(length(pl))), probability = TRUE)

# Add vertical line for VaR value
abline(v = quantile(pl, probs = (1 - alpha)),
       lty = 2, ## line type: 2 = dashed
       lwd = 2.5, ## line width: the largher, the thicker
       col = "red")
```

